---
title: "Lab: List Processing"
author: "Ryan Barent"
format: html
number-sections: true
number-depth: 2
---

::: callout
You can see the purpose of this assignment as well as the skills and knowledge you should be using and acquiring, in the [Transparency in Learning and Teaching (TILT)](tilt.qmd) document in this repository. The TILT document also contains a checklist for self-reflection that will provide some guidance on how the assignment will be graded. 
:::

# Data Source

JSON data files for this assignment were obtained from the TVMaze API for three different  Doctor Who series as well as two different spin-offs. 

- Dr. Who [2023-2025](https://www.tvmaze.com/shows/72724/doctor-who)
- Dr. Who [2005-2022](https://www.tvmaze.com/shows/210/doctor-who)
- Dr. Who [1963-1996](https://www.tvmaze.com/shows/766/doctor-who)
- [The Sarah Jane Adventures (2007-2020)](https://www.tvmaze.com/shows/970/the-sarah-jane-adventures)
- [Torchwood (2006-2011)](https://www.tvmaze.com/shows/659/torchwood)
- [Torchwood: Web of Lies (2011)](https://www.tvmaze.com/shows/26694/torchwood-web-of-lies)

# Warming Up

For this portion of the assignment, only work with the canonical Dr. Who files (drwho2023.json, drwho2005.json, drwho1963.json). 

## Parse the file

Add a code chunk that will read each of the JSON files in. 
Store the data in a `drwhoYYYY` object, where `YYYY` is the first year the series  began to air. 
How are the data objects stored?


```{r}
library(jsonlite)
library(purrr)
library(dplyr)
library(tidyverse)

test = fromJSON("drwho-766.json")

drwho1963 = read_json("drwho-766.json")
drwho2005 = read_json("drwho-210.json")
drwho2023 = read_json("drwho-72724.json")
```

---

## Examining List Data Structures

Create a nested markdown list showing what variables are nested at each level of the JSON file. Include an 'episode' object that is a stand-in for a generic episode (e.g. don't create a list with all 700+ episodes in it, just show what a single episode has). Make sure you use proper markdown formatting to ensure that the lists are rendered properly when you compile your document.

Hint: The `prettify()` function in the R package `jsonlite` will spit out a better-formatted version of a JSON file. 
```{r results="hide"}
#prettify(toJSON(drwho1963))
```

----

For Each episode there is:
- id
- url
- name
- season
- number 
- type
- airdate
- airtime
- airstamp
- runtime
- rating
- others


----

Is there any information stored in the list structure that you feel is redundant? If so, why?

Airdate and airstamp could be redundant because they both contain the date the episode was released

## Develop A Strategy

Consider what information you would need to examine the structure of Dr. Who episodes over time (show runtime, season length, specials) as well as the ratings, combining information across all three data files. 

Sketch one or more rectangular data tables that look like your expected output. Remember that if you link to an image, you must link to something with a picture extension (`.png`, `.jpg`), and if you reference a file it should be using a local path and you must also add the picture to your git repository. 


What operations will you need to perform to get the data into a form matching your sketch? Make an ordered list of steps you need to take.

---

1. map each episode as a row in dataframe
2. Identify each of variables of interest for episode (create new?)
3. combine into dataframe with all episodes across series

## Implement Your Strategy

Add a code chunk that will convert the JSON files into the table(s) you sketched above. 
Make sure that the resulting tables have the correct variable types (e.g., dates should not be stored as character variables).

Print out the first 5 rows of each table that you create (but no more)!

```{r}

drwho1963df = map_dfr(drwho1963, ~ as_tibble(map(.x, ~ if (is.null(.x)) NA else .x)))

drwho2005df = map_dfr(drwho2005, ~ as_tibble(map(.x, ~ if (is.null(.x)) NA else .x)))

drwho2023df = map_dfr(drwho2023, ~ as_tibble(map(.x, ~ if (is.null(.x)) NA else .x)))

drwho_all = bind_rows("1963" = drwho1963df,"2005" = drwho2005df,"2023" = drwho2023df,.id = "series")

head(drwho_all, 5)

```

```{r}
airdates = drwho_all %>%
  mutate(airdate = as.Date(airdate)) %>%
  group_by(series, season) %>%
  arrange(airdate) %>%
  mutate(air_diff = as.numeric(airdate - lag(airdate))) %>%
  ungroup()

```
## Examining Episode Air Dates

Visually represent the length of time between air dates of adjacent episodes within the same season, across all seasons of Dr. Who. You may need to create a factor to indicate which Dr. Who series is indicated, as there will be a Season 1 for each of the series. 
Your plot must have appropriate labels and a title.

---

```{r}
library(ggplot2)

ggplot(data = airdates) +
  geom_point(aes(x = number, y = air_diff, color = series)) +
  labs(x = "Episode Number Within a Given Season",y = "Days Since Previous Episode",title = "Differences in Episode Air Dates Across Doctor Who Series",color = "Series") 
```

---

In 2-3 sentences, explain what conclusions you might draw from the data. What patterns do you notice? Are there data quality issues?

 The primary takeaway is that the 1963 seasons lasted much longer than the other series. Most of the episodes appear to have been released at the same time as the previous or one week later. There are a few outliers 

# Timey-Wimey Series and Episodes

## Setting Up

In this section of the assignment, you will work with all of the provided JSON files. 
Use a functional programming approach to read in all of the files and bind them together. 

----

functional code goes here

----

Then, use the processing code you wrote for the previous section to perform appropriate data cleaning steps. 
At the end of the chunk, your data should be in a reasonably tidy, rectangular form with appropriate data types. 
Call this rectangular table `whoverse`. 

----

Cleaning code goes here

----


## Air Time

Investigate the air time of the episodes relative to the air date, series, and season.
It may help to know that the [watershed](https://en.wikipedia.org/wiki/Watershed_(broadcasting)) period in the UK is 9:00pm - 5:30am. 
Content that is unsuitable for minors may only be shown during this window.
What conclusions do you draw about the target audience for each show? 

How can you explain any shows in the Dr. Who universe which do not have airtimes provided?

## Another Layer of JSON

Use the show URL (`_links` > `show` > `href`) to read in the JSON file for each show. 
As with scraping, it is important to be polite and not make unnecessary server calls, so pre-process the data to ensure that you only make one server call for each show.
You should use a functional programming approach when reading in these files. 

----

Read in JSON files from URLs here

----

Process the JSON files using a functional approach and construct an appropriate table for the combined data you've acquired during this step (no need to join the data with the full `whoverse` episode-level data). 

----

Process JSON files to make a table here

----

What keys would you use to join this data with the `whoverse` episode level data? Explain.

> Explanation


## Explore!

Use the data you've assembled to answer a question you find interesting about this data.
Any graphics you make should have appropriate titles and axis labels. 
Tables should be reasonably concise (e.g. don't show all 900 episodes in a table), generated in a reproducible fashion, and formatted with markdown. 
Any results (graphics, tables, models) should be explained with at least 2-3 sentences. 

If you're stuck, consider examining the frequency of words in the episode descriptions across different series or seasons. Or, look at the episode guest cast by appending `/guestcast/` to the episode URL and see whether there are common guests across different seasons. 

----

Question goes here

----

Code goes here -- once you output a result, you should explain it using markdown text, and then start a new code chunk to continue your exploration. 

